---
title: "Cleaning Berries"
author: "Zijie Huang"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(magrittr)
library(kableExtra)

opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
```

##  Acquire and read the data

These data were collected from the USDA database selector: <a href="https://quickstats.nass.usda.gov">https://quickstats.nass.usda.gov</a>

The data were <a href="https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1">stored online</a> and then downloaded as a CSV file.


```{r}

## read the data

ag_data <- read_csv("berries.csv", col_names = TRUE)

## look at number of unique values in each column
ag_data %>% summarize_all(n_distinct) -> aa


## make a list of the columns with only one unique value
bb <- which(aa[1,]==1)

## list the 1-unique valu column names 
cn <- colnames(ag_data)[bb]
```

Data selected from the NASS database often has columns without any data or with a single repeated Values. The berries data had only 8 out of 21 columns containing meaningful data.

```{r}


## remove the 1-unique columns from the dataset
ag_data %<>% select(-all_of(bb))

aa %<>% select(-all_of(bb)) 


## State name and the State ANSI code are (sort of) redundant
## Just keep the name
ag_data %<>% select(-4)
aa %<>% select(-4) 

# ag_data$Commodity %>% unique()

# kable(head(ag_data)) %>%
#   kable_styling(font_size=12)

```


```{r}
berry <- unique(ag_data$Commodity)
nberry <- length(berry)

```

This table contains informaton about `r nberry` berries: blueberries, raspberries, and strawberries.

When the data have been cleaned and organized, the three kinds of berries will be separted into tables with the same stucture so that they can be compared.  So, working with Blueberries along demonstrates how the data will be cleaned and organized for all three kinds of berries. Only the "YEAR" time periond will be considered.

## Strawberries
```{r}
sberry <- ag_data %>% filter((Commodity=="STRAWBERRIES") & (Period=="YEAR"))
sberry %<>% select(-c(Period, Commodity))
```


```{r}
sberry %<>% separate(`Data Item`, c("B","type", "meas", "what"), sep = ", ") 

sberry %<>% separate(B,c("Red","type2"),sep=" - ")

sberry %<>% select(-Red)

sberry %<>% separate(type,c("type_left","type_right"),sep = " - ")

sberry[is.na(sberry)] <- " "

sberry %<>% mutate(type = "",production = "",Avg = "", Measures = "", Materials = "", Chemicals = "")


# Solve each unique 
## "STRAWBERRIES - ACRES HARVESTED"
sberry1 <- sberry %>% 
  filter(type2 == "ACRES HARVESTED") %>%
  mutate(production = "ACRES HARVESTED")

## "STRAWBERRIES - ACRES PLANTED" 
sberry2 <- sberry %>%
  filter(type2 == "ACRES PLANTED") %>%
  mutate(production = "ACRES PLANTED")

## "STRAWBERRIES - PRODUCTION, MEASURED IN $" 
##& "STRAWBERRIES - PRODUCTION, MEASURED IN CWT"
##& "STRAWBERRIES - YIELD, MEASURED IN CWT / ACRE"
sberry3 <- sberry %>%
  filter(type2 %in% c("PRODUCTION","YIELD")) %>%
  mutate(production = type2,
         Measures = type_left)

## "STRAWBERRIES, BEARING"
sberry4 <- sberry %>%
  filter(type_left == "BEARING") %>%
  mutate(type = type_left,
         production = type_right,
         Measures = meas,
         Avg = what)

## "STRAWBERRIES, FRESH MARKET - PRODUCTION"
sberry5 <- sberry %>%
  filter(type_left == "FRESH MARKET",
         type_right == "PRODUCTION") %>%
  mutate(production = paste(type_left,type_right,sep = " - "),
         Measures = meas)

## ""STRAWBERRIES, FRESH MARKET, UTILIZED - PRODUCTION"
sberry6 <- sberry %>%
  filter(meas == "UTILIZED - PRODUCTION") %>%
  mutate(production = paste(type_left,meas,sep = " - "),
         Measures = what)

## STRAWBERRIES, PROCESSING - PRODUCTION"
sberry7 <- sberry %>%
  filter(type_right == "PRODUCTION",
         type_left %in% c("NOT SOLD","PROCESSING","UTILIZED")) %>%
  mutate(production = paste(type_left,type_right,sep = " - "),
         Measures = meas)

# Combine together
sberry_clean <- bind_rows(list(sberry1,sberry2,sberry3,sberry4,sberry5,sberry6,sberry7))

# drop columns
sberry_clean %<>% select(-c(type2,type_left,type_right,meas,what))

# Clean Domain and Domain Category
sberry_clean %<>% separate(Domain,c("Domain_left","Domain_right"),sep = ", ")

sberry_clean %<>% separate(`Domain Category`,c("Chem","right"),sep = ", ")

sberry_clean %<>% separate(Chem,c("Chem_1","Chem_2"),sep = ": ")

sberry_clean %<>% separate(right,c("right_1","right_2"),sep = ": ")

sberry_clean[is.na(sberry_clean)] <- " "

## "TOTAL"
sberry_clean1 <- sberry_clean %>%
  filter(Domain_left == "TOTAL") %>%
  mutate(Chemicals = Domain_left,
         Materials = Chem_1)

## "Chem"
sberry_clean2 <- sberry_clean %>%
  filter(Domain_left == "CHEMICAL") %>%
  mutate(Chemicals = right_1,
         Materials = right_2)

## "Fertilizer"
sberry_clean3 <- sberry_clean %>%
  filter(Domain_left == "FERTILIZER") %>%
  mutate(Chemicals = paste(Chem_2,Domain_left,sep = " "))

# Combine together
sberry_clean <- bind_rows(list(sberry_clean1,sberry_clean2,sberry_clean3))

# arrange columns
sberry_clean %<>% select(-c(Domain_left,Domain_right,Chem_1,Chem_2,right_1,right_2))

sberry_clean <- sberry_clean %>% select(-Value,Value)

sberry_clean %<>% arrange(desc(Year),State)

# Check
# kable(head(sberry_clean,n=10)) %>%
#   kable_styling(font_size=12)

# Export data
# write_csv(sberry_clean,"rberry_clean.csv")

```